---
title: "candiate_gene_analysis"
author: "Sylvain Moser"
date: "2/11/2021"
output: 
  html_document:
    code_folding: hide
params:
  candidate_analysis_exposure_results: !r NULL
  exposure_final_processed_results: !r NULL
  candidate_analysis_therapy_results: !r NULL
  therapy_final_processed_results: !r NULL
  candidate_analysis_expression_therapy_results: !r NULL
  candidate_analysis_expression_exposure_results: !r NULL
  candidate_analysis_iurato_therapy_results: !r NULL
  candidate_analysis_iurato_exposure_results: !r NULL
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The candidate gene analysis was performed using a separate script and results are depicted here

```{r}
library(dplyr)
```

## Exposure

```{r}
candidates=read.table(params$candidate_analysis_exposure_results,h=T,stringsAsFactors = F)

load(params$exposure_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```



### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```

## Therapy

```{r}
candidates=read.table(params$candidate_analysis_therapy_results,h=T,stringsAsFactors = F)

load(params$therapy_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```

### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```

### Candidate gene analysis based on the significant genes expression changes during CBT in Martins et al : 

#### Exposure

```{r}
candidates=read.table(params$candidate_analysis_expression_exposure_results,h=T,stringsAsFactors = F)
load(params$exposure_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```



### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```

#### Therapy

```{r}
candidates=read.table(params$candidate_analysis_expression_therapy_results,h=T,stringsAsFactors = F)
load(params$therapy_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```

### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```

### Candidate gene analysis based on the significant CpGs in the case control EWAS of Iurato (and two Cpgs from the Naik paper : 

#### Exposure

```{r}
candidates=read.table(params$candidate_analysis_iurato_exposure_results,h=T,stringsAsFactors = F)
load(params$exposure_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```



### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```



#### Therapy

```{r}
candidates=read.table(params$candidate_analysis_iurato_therapy_results,h=T,stringsAsFactors = F)
load(params$therapy_final_processed_results)
```

```{r}
print (head(candidates %>% arrange(qval)))
```

### checking if distribution of pvalues in the candidates genes is different as for all other CpGs: 

Using a logistic regression with the binary candidates variable and p-value as a predictor: 

```{r}
log_df=final_processed_results %>% select(island,best_lmm_pval) %>% mutate(candidate=ifelse(island %in% candidates$island,1,0))
log_mod=glm(candidate ~ best_lmm_pval,data=log_df,family = "binomial")
summary(log_mod)
```