---
title: "KLF16_methylation_CBT_profile"
author: "Sylvain Moser"
date: "2/2/2021"
output: 
  html_document:
    code_folding: hide
params:
  methylation_df: !r NULL
  pheno: !r NULL
  methylation_residuals: !r NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
prefix_statgen=ifelse(file.exists("/psycl/g/mpsstatgen/symo/dummy_file.txt"),"/psycl/g/mpsstatgen/symo/anxiety_methylation/","/Users/sylvain_moser/psycl_statgen/symo/anxiety_methylation/")
```


```{r}
load(params$methylation_df)

exposure_ID=meth_df %>% filter(Timepoint %in% c("b1","pe1","p24_1")) %>% pull(NID)
therapy_ID=meth_df %>% filter(Timepoint %in% c("T0","T4","E","K1")) %>% pull(NID)
```

# Profiles using raw methylation data 

## whole CBT profile: 

```{r}
time_df=meth_df %>% select(NID,Timepoint,cg10761804) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))


plot_KLF16_CBT=ggplot(time_df,aes(x=Time,y=cg10761804,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 during Therapy",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_KLF16_CBT
```

## Only exposure and following timepoints: 
```{r}
time_df=meth_df %>% select(NID,Timepoint,cg10761804) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_KLF16_postint=ggplot(time_df,aes(x=Time,y=cg10761804,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),,axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_KLF16_postint
```

### color coded for the remission status: 
```{r}
pheno=read.csv(params$pheno,sep=";")
pheno_df=pheno %>%
  mutate(response=case_when(HAMAend/HAMA <=0.5 ~ 1,HAMAend/HAMA > 0.5 ~ 0)) %>% 
  mutate(remission=case_when(HAMAend <=7 ~1,HAMAend >7 ~0)) %>% 
  select(c(NID,response,remission))

remission_df=left_join(time_df,pheno_df)

plot_KLF16_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16",y="beta values",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_KLF16_postint_remission
```

### color coded for the response status: 

```{r}
plot_KLF16_postint_response=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16",y="beta values",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_KLF16_postint_response

rm(meth_df,remission_df,time_df)
```

# Profiles using residuals of the methylation data after regressing the immune cell types

```{r}
meth_residuals=readRDS(params$methylation_residuals)

```

## whole CBT profile: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg10761804) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
  ))


plot_KLF16_CBT_residuals=ggplot(time_df,aes(x=Time,y=cg10761804,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 during therapy",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_KLF16_CBT_residuals
```

### color coded for the remission status: 
```{r}
remission_df=left_join(time_df,pheno_df)

plot_KLF16_CBT_residuals_remission=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_KLF16_CBT_residuals_remission
```

### color coded for the response status: 

```{r}
plot_KLF16_CBT_residuals_response=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
  plot_KLF16_CBT_residuals_response
```

## only the exposure and following time points: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg10761804) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_KLF16_residuals_postint=ggplot(time_df,aes(x=Time,y=cg10761804,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_KLF16_residuals_postint
```

### color coded for the remission status: 

```{r}
remission_df=left_join(time_df,pheno_df)

plot_KLF16_CBT_residuals_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_KLF16_CBT_residuals_postint_remission
```
## color coded for the response status: 

```{r}
plot_KLF16_CBT_residuals_postint_response=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_KLF16_CBT_residuals_postint_response
```

# Profiles using residuals of the methylation data after regressing the immune cell types centered at 0 for BE:

```{r}
centered_time_df=time_df %>% group_by(NID) %>% group_map(~mutate(.,cg10761804=cg10761804  - as.numeric(.$cg10761804[.$Timepoint=="b1"])),.keep=T) %>% bind_rows()
remission_df=left_join(centered_time_df,pheno_df)
```

## remission

```{r}
plot_KLF16_CBT_residuals_centered_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="delta resid. methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_KLF16_CBT_residuals_centered_postint_remission
```
## statistical test for the difference between remission0 and remission1 group: 

```{r}
t.test(remission_df %>% dplyr::filter(Time==1 & remission==0) %>% pull (cg10761804),remission_df %>% dplyr::filter(Time==1 & remission==1) %>% pull (cg10761804))
t.test(remission_df %>% dplyr::filter(Time==2 & remission==0) %>% pull (cg10761804),remission_df %>% dplyr::filter(Time==2 & remission==1) %>% pull (cg10761804))
t.test(remission_df %>% dplyr::filter(Time==3 & remission==0) %>% pull (cg10761804),remission_df %>% dplyr::filter(Time==3 & remission==1) %>% pull (cg10761804))
```

## response

```{r}
plot_KLF16_CBT_residuals_centered_postint_response=ggplot(remission_df,aes(x=Time,y=cg10761804,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "KLF16 after Exposure",y="delta resid. methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_KLF16_CBT_residuals_centered_postint_response
```

