---
title: "MAD1L1_methylation_CBT_profile"
author: "Sylvain Moser"
date: "3/8/2021"
output: 
  html_document:
    code_folding: hide
params:
  methylation_df: !r NULL
  pheno: !r NULL
  methylation_residuals: !r NULL

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
prefix_statgen=ifelse(file.exists("/psycl/g/mpsstatgen/symo/dummy_file.txt"),"/psycl/g/mpsstatgen/symo/anxiety_methylation/","/Users/sylvain_moser/psycl_statgen/symo/anxiety_methylation/")
```


```{r}
load(params$methylation_df)
exposure_ID=meth_df %>% filter(Timepoint %in% c("b1","pe1","p24_1")) %>% pull(NID)
therapy_ID=meth_df %>% filter(Timepoint %in% c("T0","T4","E","K1")) %>% pull(NID)
```

# cg13815311 from the exposure analysis


# Profiles using raw methylation data 

## whole CBT profile: 

```{r}
time_df=meth_df %>% select(NID,Timepoint,cg13815311) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))


plot_MAD1L1_CBT=ggplot(time_df,aes(x=Time,y=cg13815311,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during Therapy",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT
```

## Only exposure and following timepoints: 
```{r}
time_df=meth_df %>% select(NID,Timepoint,cg13815311) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_postint=ggplot(time_df,aes(x=Time,y=cg13815311,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),,axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_postint
```

### color coded for the remission status: 
```{r}
pheno=read.csv(params$pheno,sep=";")
pheno_df=pheno %>%
  mutate(response=case_when(HAMAend/HAMA <=0.5 ~ 1,HAMAend/HAMA > 0.5 ~ 0)) %>% 
  mutate(remission=case_when(HAMAend <=7 ~1,HAMAend >7 ~0)) %>% 
  select(c(NID,response,remission))

remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_postint_remission=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_postint_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_postint_response=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_postint_response

rm(remission_df,time_df)
```

# Profiles using residuals of the methylation data after regressing the immune cell types

```{r}
meth_residuals=readRDS(params$methylation_residuals)
```

## whole CBT profile: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg13815311) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
  ))


plot_MAD1L1_CBT_residuals=ggplot(time_df,aes(x=Time,y=cg13815311,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during therapy",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT_residuals
```

### color coded for the remission status: 
```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_remission=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_response=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
  plot_MAD1L1_CBT_residuals_response
```

## only the exposure and following time points: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg13815311) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_residuals_postint=ggplot(time_df,aes(x=Time,y=cg13815311,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_residuals_postint
```

### color coded for the remission status: 

```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_postint_remission=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_postint_remission
```
## color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_postint_response=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_postint_response
```

# Profiles using residuals of the methylation data after regressing the immune cell types centered at 0 for BE:

```{r}
centered_time_df=time_df %>% group_by(NID) %>% group_map(~mutate(.,cg13815311=cg13815311  - as.numeric(.$cg13815311[.$Timepoint=="b1"])),.keep=T) %>% bind_rows()
remission_df=left_join(centered_time_df,pheno_df)
```

## remission

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_remission=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_centered_postint_remission
```
## statistical test for the difference between remission0 and remission1 group: 

```{r}
t.test(remission_df %>% dplyr::filter(Time==1 & remission==0) %>% pull (cg13815311),remission_df %>% dplyr::filter(Time==1 & remission==1) %>% pull (cg13815311))
t.test(remission_df %>% dplyr::filter(Time==2 & remission==0) %>% pull (cg13815311),remission_df %>% dplyr::filter(Time==2 & remission==1) %>% pull (cg13815311))
t.test(remission_df %>% dplyr::filter(Time==3 & remission==0) %>% pull (cg13815311),remission_df %>% dplyr::filter(Time==3 & remission==1) %>% pull (cg13815311))
```

## response

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_response=ggplot(remission_df,aes(x=Time,y=cg13815311,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_centered_postint_response
```

# cg10418812 from the therapy candidate analysis

## whole CBT profile: 

```{r}
time_df=meth_df %>% select(NID,Timepoint,cg10418812) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))


plot_MAD1L1_CBT=ggplot(time_df,aes(x=Time,y=cg10418812,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during Therapy",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT
```

## Only exposure and following timepoints: 
```{r}
time_df=meth_df %>% select(NID,Timepoint,cg10418812) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_postint=ggplot(time_df,aes(x=Time,y=cg10418812,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),,axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_postint
```

### color coded for the remission status: 
```{r}
pheno=read.csv(params$pheno,sep=";")
pheno_df=pheno %>%
  mutate(response=case_when(HAMAend/HAMA <=0.5 ~ 1,HAMAend/HAMA > 0.5 ~ 0)) %>% 
  mutate(remission=case_when(HAMAend <=7 ~1,HAMAend >7 ~0)) %>% 
  select(c(NID,response,remission))

remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_postint_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_postint_response=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_postint_response

rm(remission_df,time_df)
```

# Profiles using residuals of the methylation data after regressing the immune cell types

```{r}
meth_residuals=readRDS(params$methylation_residuals)
```

## whole CBT profile: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg10418812) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
  ))


plot_MAD1L1_CBT_residuals=ggplot(time_df,aes(x=Time,y=cg10418812,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during therapy",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT_residuals
```

### color coded for the remission status: 
```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_remission=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_response=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
  plot_MAD1L1_CBT_residuals_response
```

## only the exposure and following time points: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg10418812) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_residuals_postint=ggplot(time_df,aes(x=Time,y=cg10418812,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_residuals_postint
```

### color coded for the remission status: 

```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_postint_remission
```
## color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_postint_response=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_postint_response
```

# Profiles using residuals of the methylation data after regressing the immune cell types centered at 0 for BE:

```{r}
centered_time_df=time_df %>% group_by(NID) %>% group_map(~mutate(.,cg10418812=cg10418812  - as.numeric(.$cg10418812[.$Timepoint=="b1"])),.keep=T) %>% bind_rows()
remission_df=left_join(centered_time_df,pheno_df)
```

## remission

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_remission=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_centered_postint_remission
```
## statistical test for the difference between remission0 and remission1 group: 

```{r}
t.test(remission_df %>% dplyr::filter(Time==1 & remission==0) %>% pull (cg10418812),remission_df %>% dplyr::filter(Time==1 & remission==1) %>% pull (cg10418812))
t.test(remission_df %>% dplyr::filter(Time==2 & remission==0) %>% pull (cg10418812),remission_df %>% dplyr::filter(Time==2 & remission==1) %>% pull (cg10418812))
t.test(remission_df %>% dplyr::filter(Time==3 & remission==0) %>% pull (cg10418812),remission_df %>% dplyr::filter(Time==3 & remission==1) %>% pull (cg10418812))
```

## response

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_response=ggplot(remission_df,aes(x=Time,y=cg10418812,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_centered_postint_response
```

# cg24577389 from the therapy candidate analysis

## whole CBT profile: 

```{r}
time_df=meth_df %>% select(NID,Timepoint,cg24577389) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))


plot_MAD1L1_CBT=ggplot(time_df,aes(x=Time,y=cg24577389,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during Therapy",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT
```

## Only exposure and following timepoints: 
```{r}
time_df=meth_df %>% select(NID,Timepoint,cg24577389) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_postint=ggplot(time_df,aes(x=Time,y=cg24577389,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="beta values")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),,axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_postint
```

### color coded for the remission status: 
```{r}
pheno=read.csv(params$pheno,sep=";")
pheno_df=pheno %>%
  mutate(response=case_when(HAMAend/HAMA <=0.5 ~ 1,HAMAend/HAMA > 0.5 ~ 0)) %>% 
  mutate(remission=case_when(HAMAend <=7 ~1,HAMAend >7 ~0)) %>% 
  select(c(NID,response,remission))

remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_postint_remission=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_postint_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_postint_response=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="beta values",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_postint_response

rm(remission_df,time_df)
```

# Profiles using residuals of the methylation data after regressing the immune cell types

```{r}
meth_residuals=readRDS(params$methylation_residuals)
```

## whole CBT profile: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg24577389) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
  ))


plot_MAD1L1_CBT_residuals=ggplot(time_df,aes(x=Time,y=cg24577389,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 during therapy",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_CBT_residuals
```

### color coded for the remission status: 
```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_remission=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_remission
```

### color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_response=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4,5,6),labels=c("T0","T4","BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
  plot_MAD1L1_CBT_residuals_response
```

## only the exposure and following time points: 

```{r}
time_df=meth_residuals %>% select(NID,Timepoint,cg24577389) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="b1" ~ 0,
    Timepoint=="pe1" ~ 1,
    Timepoint=="p24_1" ~ 2,
    Timepoint=="E" ~ 3,
    Timepoint=="K1" ~ 4,
  ))

plot_MAD1L1_residuals_postint=ggplot(time_df,aes(x=Time,y=cg24577389,group=NID,colour=NID))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.")+
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),axis.line = element_line(colour ="darkgrey"))+
stat_summary(fun=mean,geom="line",lwd=2,aes(group=1))
plot_MAD1L1_residuals_postint
```

### color coded for the remission status: 

```{r}
remission_df=left_join(time_df,pheno_df)

plot_MAD1L1_CBT_residuals_postint_remission=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_postint_remission
```
## color coded for the response status: 

```{r}
plot_MAD1L1_CBT_residuals_postint_response=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="residuals methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_postint_response
```

# Profiles using residuals of the methylation data after regressing the immune cell types centered at 0 for BE:

```{r}
centered_time_df=time_df %>% group_by(NID) %>% group_map(~mutate(.,cg24577389=cg24577389  - as.numeric(.$cg24577389[.$Timepoint=="b1"])),.keep=T) %>% bind_rows()
remission_df=left_join(centered_time_df,pheno_df)
```

## remission

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_remission=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(remission)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="remission")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(remission)))
plot_MAD1L1_CBT_residuals_centered_postint_remission
```
## statistical test for the difference between remission0 and remission1 group: 

```{r}
t.test(remission_df %>% dplyr::filter(Time==1 & remission==0) %>% pull (cg24577389),remission_df %>% dplyr::filter(Time==1 & remission==1) %>% pull (cg24577389))
t.test(remission_df %>% dplyr::filter(Time==2 & remission==0) %>% pull (cg24577389),remission_df %>% dplyr::filter(Time==2 & remission==1) %>% pull (cg24577389))
t.test(remission_df %>% dplyr::filter(Time==3 & remission==0) %>% pull (cg24577389),remission_df %>% dplyr::filter(Time==3 & remission==1) %>% pull (cg24577389))
```

## response

```{r}
plot_MAD1L1_CBT_residuals_centered_postint_response=ggplot(remission_df,aes(x=Time,y=cg24577389,group=NID,colour=as.factor(response)))+geom_line()+
  scale_x_continuous(name="Timepoints",breaks=c(0,1,2,3,4),labels=c("BE","PE1h","PE24h","E","K"))+
  labs(title = "MAD1L1 after Exposure",y="delta resid. methyl.",colour="response")+
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_blank(),text = element_text(size = 30),axis.ticks=element_blank(),legend.title=element_text(size = 20),legend.text = element_text(size = 20),axis.line = element_line(colour ="darkgrey"))+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=as.factor(response)))
plot_MAD1L1_CBT_residuals_centered_postint_response
```

## linear model to test for regulation over all time points: 
### using the raw data 

```{r}
time_df=meth_df %>% select(NID,Timepoint,cg10418812,cg24577389,CD8T,CD4T,NK,Bcell,Mono,Gran,age,Sex,Zigis) %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))
```

```{r}
library(lmerTest)
lmm_2_full=lmer(cg10418812 ~ Time +I(Time^2) + (1|NID) +CD8T +CD4T + NK+ Bcell + Mono+ Gran+ age + Sex + Zigis,data=time_df,REML = TRUE)
anova(lmm_2_full)
```
```{r}
lmm_2_full=lmer(cg24577389 ~ Time +I(Time^2) + (1|NID) +CD8T +CD4T + NK+ Bcell + Mono+ Gran+ age + Sex + Zigis,data=time_df,REML = TRUE)
anova(lmm_2_full)
```

### using the residuals:

```{r}
meth_residuals=left_join(meth_residuals %>% select(NID,Timepoint,cg10418812,cg24577389),meth_df %>% select(NID,Timepoint,Sex,age,Zigis))
time_df=meth_residuals %>% filter (NID %in% intersect(exposure_ID,therapy_ID)) %>% 
  mutate(Time=case_when(
    Timepoint=="T0" ~ 0,
    Timepoint=="T4" ~ 1,
    Timepoint=="b1" ~ 2,
    Timepoint=="pe1" ~ 3,
    Timepoint=="p24_1" ~ 4,
    Timepoint=="E" ~ 5,
    Timepoint=="K1" ~ 6,
))
```

```{r}
lmm_2_full=lmer(cg10418812 ~ Time +I(Time^2) + (1|NID)+ Sex + age + Zigis,data=time_df,REML = TRUE)
anova(lmm_2_full)
```

```{r}
lmm_2_full=lmer(cg24577389 ~ Time +I(Time^2) + (1|NID)+ Sex + age + Zigis,data=time_df,REML = TRUE)
anova(lmm_2_full)
```